Periodic Traveling Waves in an Integro-Difference Equation With a Nonmonotone Growth Function and Strong Allee Effect

### 2021-06-14 paper_draft3.tex
continued edits

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{rem}[thm]{Remark}
\newtheorem{hyp}[thm]{Hypothesis}
\newtheorem{ex}[thm]{Example}

@@

Under these assumptions, the cumulative density function $F(x)=\int_{-\infty}^{x}k(y)\,dy$ has a well-defined right inverse $F^{-1}$ satisfying $F(F^{-1}(p))=p$ for all $0<p<1$. Furthermore, it satisfies the symmetry relation $F(-x)=1-F(x)$.

 Let $w_1$ and $w_2$ be two functions belonging to $C(\mathbb R)$ defined by
\begin{equation} \label{w1}
w_1(x) = \int_x^\infty k(y) \, dy
\end{equation}
and
\begin{equation} \label{w2}
w_2(x) = \int_{-\infty}^{\infty} k(y) g(w_1(x-y)) \, dy 
\end{equation}

@@

\begin{proof} Let $x_\a=\sup\{x\in\mathbb R\mid w_1(x)\geq\a\}$ and $x_\b=\sup\{x\in\mathbb R\mid w_1(x)\geq\b\}$. It follows that $w_1(x)>\b$ for $x<x_\b$, $\a\leq w_1(x)\leq\b$ for $x_\b\leq x\leq x_\a$, and $w_1(x)<\a$ for $x>x_\a$. Thus, composing $w_1(x)$ with the growth function $g$ yields
\begin{equation} \label{gw1}
g(w_1(x)) = \begin{cases}
\mu & x < x_\beta \\
1 & x_\beta \leq x \leq x_\a \\
0 & x > x_\a
\end{cases} \end{equation}

Applying the convolution operator yields a formula for $w_2(x)$:
\begin{equation} \label{w2calculation}
\begin{aligned}
w_2(x) &= \int_{-\infty}^{\infty} k(y) g(w_1(x-y)) \, dy \\
&= \int_{x-x_\a}^{x-x_\b} k(y)\,dy + \m \int_{x-\beta}^{\infty} k(y)\,dy \\
&= \int_{x-x_\a}^{\infty} k(y)\,dy - (1-\m) \int_{x-x_\b}^{\infty} k(y)\,dy
\end{aligned}
\end{equation}

@@

It follows that $w_2(x)\geq\a$ for $x\leq 2c^*$ and $w_2(x)<\alpha$ for $x>2c^*$. Along with the assumption that $\sup_{x\in\mathbb R}w_2(x)\leq\b$, we can compose with $g$ as follows:
\begin{equation} \label{gw2}
g(w_2(x)) = \begin{cases}
1 & x \leq 2c^* \\
0 & x > 2c^*
\end{cases} \end{equation}

Taking convolution with $k$, we obtain
\begin{equation}
Q[w_2](x) = \int_{x-x_0}^{\infty} k(y)\,dy = w_1(x-2c^*).
\end{equation}
%By defining $c^* = \frac{1}{2}x_0$, and applying the integro-difference operator, we can see that $w_1$ and $w_2$ are both solutions to the periodic traveling wave equation with mean speed $c^*$:
%\begin{equation}
%Q^2[w_1](x) = w_1(x-2c^*); \quad Q^2[w_2](x) = w_2(x-2c^*)
%\end{equation}
\end{proof}

\begin{thm} \label{theorem1}
If $\sup_{x\in\mathbb R}w_2(x)\leq \b$, then the recurrence \eqref{q} has a periodic traveling wave solution given by
\begin{equation} \label{ptw}
W_n(x) = \begin{cases}
w_1(x-nc^*) & n \text{ even} \\
w_2(x-nc^*) & n \text{ odd}
\end{cases}
\end{equation}
\end{thm}

\begin{proof}
DO THIS
\end{proof}

\begin{thm}  \label{theorem2}
For all $\epsilon>0$, the sequence $(u_n)_{n=0}^{\infty}$ defined in Theorem \ref{theorem1} satisfies
\begin{equation}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n}(x)=1
\end{equation}
and
\begin{equation}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n+1}(x)=m
\end{equation}
and 
\begin{equation}
\lim_{n\to\infty}\sup_{x>n(c^*+\epsilon)}u_n(x)=0.
\end{equation}
\end{thm}

@@

\begin{thm} Suppose $||w_2||_\infty \leq b$ and $k(x)$ has compact support. Assume the following hold:

\begin{enumerate}[i.)]
\item $c^*$ is positive,\
\item the initial condition $u_0$ has compact support;
\item the set $A=\{x\in\mathbb R:a\leq u_0(x)\leq b\}$  is connected and sufficiently large.
\end{enumerate}
 Then the sequence $(u_n)_{n=0}^{\infty}$ defined by $u_{n+1}=Q[u_n]$, $n\geq 0$, satisfies
\begin{equation}
\lim_{n\to\infty}\inf_{|x|<2nc}u_{2n}(x)=1, \quad \forall c\in(0,c^*)
\end{equation}
and
\begin{equation}
\lim_{n\to\infty}\inf_{|x|<2nc}u_{2n+1}(x)=m, \quad \forall c\in(0,c^*)
\end{equation}
and 
\begin{equation}
\lim_{n\to\infty}\sup_{|x|>nc}u_n(x)=0, \quad \forall c\in(c^*,\infty).
\end{equation}
\end{thm}

\begin{proof}
Let $(-\sigma,\sigma)$ be the support of $k$, for $0<\sigma<\infty$. Since $A$ is connected, we may also assume by the translation invariance property that $A=[-r,r]$ for some $r \geq 0$. Furthermore, since $u_0$ has compact support we must have $u_0(x)=0$ outside a bounded interval $(r^-,r^+) \supset A$, where $r^- < -r \leq r < r^+$.

@@

\section{Example calculations}

>>>>

%\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ex}[thm]{Example}

@@

\begin{prop} \label{prop1}
Let $\mathcal B$ denote the space of bounded, continuous, non-negative functions $\mathbb R \to \mathbb R$. $Q$ is a bounded nonlinear operator $\mathcal B \to \mathcal B$, and satisfies the following properties for all $u \in \mathcal B$:

\begin{enumerate}[(i.)]
\item \it{(Translation invariance)} $Q[T^t[u]]=T^t[Q[u]]$ for any translation operator $T^t[u](x)=u(x+t)$.
\item \it{(Limit-preserving)} If $u(x)\to\ell$ as $x\to\infty$, then $Q[u](x)\to g(\ell)$ as $x\to\infty$; likewise as $x\to-\infty$.
\item \it{(Symmetry-preserving)} If $u(x-a)=u(x+a)$ then $Q[u](x-a)=Q[u](x+a)$.
\end{enumerate}
\end{prop}

 the cumulative density function $F(x)=\int_{-\infty}^{x}k(y)\,dy$ has a well-defined right inverse $F^{-1}$ satisfying $F(F^{-1}(p))=p$ for all $0<p<1$. Furthermore, it satisfies the symmetry relation $F(-x)=1-F(x)$.

 Let $w_1$ and $w_2$ be two functions belonging to $C(\mathbb R)$ defined by
\begin{equation} \label{w1}
w_1(x) := \int_x^\infty k(y) \, dy
\end{equation}
and
\begin{equation} \label{w2}
w_2(x) := Q[w_1](x) = \int_{-\infty}^{\infty} k(y) g(w_1(x-y)) \, dy 
\end{equation}

@@

\begin{proof} Let $x_\a=\sup\{x\in\mathbb R\mid w_1(x)\geq\a\}$ and $x_\b=\sup\{x\in\mathbb R\mid w_1(x)\geq\b\}$. It follows that $w_1(x)>\b$ for $x<x_\b$, $\a\leq w_1(x)\leq\b$ for $x_\b\leq x\leq x_\a$, and $w_1(x)<\a$ for $x>x_\a$. Applying the integro-difference operator to $w_1$ yields
\begin{equation} \label{qw1calculation}
\begin{aligned}
Q[w_1](x) &= \int_{-\infty}^{\infty} k(x-y) g(w_1(y)) \, dy \\
&= \m \int_{-\infty}^{x_\b} k(x-y) \, dy + \int_{x_\b}^{x_\a} k(x-y) \, dy \\
&= \m \int_{x-x_\b}^{\infty} k(y) \, dy + \int_{x-x_\a}^{x-x_\b} k(y) \, dy \\
&= \int_{x-x_\a}^{\infty} k(y) \, dy - (1-\m) \int_{x-x_\b}^{\infty} k(y) \, dy
\end{aligned} \end{equation}

@@

It follows that $w_2(x)\geq\a$ for $x\leq 2c^*$ and $w_2(x)<\alpha$ for $x>2c^*$. Along with the assumption that $\sup_{x\in\mathbb R}w_2(x)\leq\b$, we can apply the integro-difference operator once more:
\begin{equation} \begin{aligned}
Q[w_2](x) %&= \int_{-\infty}^{\infty} k(x-y) g(w_2(y)) \, dy \\
&= \int_{-\infty}^{2c^*} k(x-y) \, dy \\
&= \int_{x-2c^*}^{\infty} k(y) \, dy &= w_1(x-2c^*).
\end{aligned} \end{equation}
%By defining $c^* = \frac{1}{2}x_0$, and applying the integro-difference operator, we can see that $w_1$ and $w_2$ are both solutions to the periodic traveling wave equation with mean speed $c^*$:
%\begin{equation}
%Q^2[w_1](x) = w_1(x-2c^*); \quad Q^2[w_2](x) = w_2(x-2c^*)
%\end{equation}
\end{proof}

\begin{thm} \label{theorem1}
If $\sup_{x\in\mathbb R}w_2(x)\leq \b$, then the recurrence \eqref{q} has a periodic traveling wave solution given by
\begin{equation} \label{ptw}
u_n(x) = \begin{cases}
w_1(x-nc^*) & n \text{ even} \\
w_2(x-nc^*) & n \text{ odd}
\end{cases}
\end{equation}
\end{thm}

\begin{proof}
It suffices to show that $u_{n+1}=Q[u_n]$ for all $n\geq 0$; this can be done via induction. The basis case is
\begin{equation}
u_1(x) = w_2(x) = Q[w_1](x) = Q[u_0](x)
\end{equation}
For the inductive step, suppose $u_{n+1}=Q[u_n]$ for some $n\geq 0$. We can proceed in cases. If $n$ is even, then $u_n(x)=w_1(x-nc^*)$ and $u_{n+1}(x)=w_2(x-nc^*)$; we have
\begin{equation}
u_{n+1}(x) = w_2(x-nc^*) = Q[w_1](x-nc^*) = Q[u_n](x)
\end{equation}
If $n$ is odd, then $u_n(x)=w_2(x-nc^*)$, and $u_{n+1}(x)=w_1(x-(n+1)c^*)$. By Lemma \ref{lemma1}, we have
\begin{equation}
u_{n+1}(x) = w_1(x-(n+1)c^*) = Q[w_2](x-nc^*) = Q[u_n](x)
\end{equation}
This completes the proof.
\end{proof}

\begin{thm}  \label{theorem2}
If $\sup_{x\in\mathbb R}w_2(x)\leq \b$, then for all $\epsilon>0$, the sequence \ref{ptw} satisfies
\begin{equation}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n}(x)=1
\end{equation}
and
\begin{equation}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n+1}(x)=m
\end{equation}
and 
\begin{equation}
\lim_{n\to\infty}\sup_{x>n(c^*+\epsilon)}u_n(x)=0.
\end{equation}
\end{thm}

\begin{proof}
\begin{equation} \begin{aligned}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n}(x)
&= \lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}w_1(x-2nc^*) \\
&= \lim_{n\to\infty}\inf_{x<-2n\epsilon}w_1(x) \\
&= \liminf_{x\to-\infty}w_1(x) &= 1.
\end{aligned} \end{equation}

\begin{equation} \begin{aligned}
\lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}u_{2n+1}(x)
&= \lim_{n\to\infty}\inf_{x<2n(c^*-\epsilon)}w_2(x-2nc^*) \\
&= \lim_{n\to\infty}\inf_{x<-2n\epsilon}w_2(x) \\
&= \liminf_{x\to-\infty}w_2(x) &= m.
\end{aligned} \end{equation}

\begin{equation} \begin{aligned}
\lim_{n\to\infty} \sup_{x>n(c^*+\epsilon)} u_n(x) 
&\leq \lim_{n\to\infty} \sup_{x>n(c^*+\epsilon)} \max\{w_1(x-nc^*),w_2(x-nc^*)\} \\
&= \lim_{n\to\infty} \sup_{x>n\epsilon} \max\{w_1(x),w_2(x)\} \\
&= \limsup_{x\to\infty} \max\{w_1(x),w_2(x)\} &= 0.
\end{aligned} \end{equation}
Since the left hand side must also be non-negative, equality follows.
\end{proof}

@@

\begin{thm} Let $u_0(x)$ be a non-negative continuous function with compact support such that the set $A=\{x\in\mathbb R:\a\leq u_0(x)\leq \b\}$ is connected. Suppose $k(x)$ has compact support, $\sup_{x\in\mathbb R}w_2(x)\leq \b$, and $c^* > 0$. Then, for $A$ sufficiently large, the sequence $(u_n)_{n=0}^{\infty}$ defined by $u_{n+1}=Q[u_n]$, $n\geq 0$ spreads with speed $c^*$, i.e., 
\begin{equation}
\lim_{n\to\infty}\inf_{|x|<n(c^*-\epsilon)}u_{n}(x)>0,
\end{equation}
and 
\begin{equation}
\lim_{n\to\infty}\sup_{|x|>n(c^*+\epsilon)}u_n(x)=0.
\end{equation}
for all $\epsilon>0$.
\end{thm}

\begin{proof}
By the translation invariance property of $Q$, we may assume without loss of generality $A=[-r,r]$, for some $r>0$. It follows that $u_0(x)<\a$ for $|x|>r$ and $\a\leq u_0(x)\leq\b$ for $|x|$Furthermore, there exists $\sigma>0$ such that $k(x)>0$ if and only if $|x|<\sigma$.

Let $(-\sigma,\sigma)$ be the support of $k$, for $0<\sigma<\infty$. Since $A$ is connected, we may also assume by the translation invariance property that $A=[-r,r]$ for some $r \geq 0$. Furthermore, since $u_0$ has compact support we must have $u_0(x)=0$ outside a bounded interval $(r^-,r^+) \supset A$, where $r^- < -r \leq r < r^+$.

@@

\section{Appendix}

\begin{proof}[Proof of Proposition \ref{prop1}]

Let $u(x)$ be non-negative, continuous, and bounded.

\begin{enumerate}[(i.)]
\item For $t\in\mathbb R$, let $T^t$ denote the translation operator defined $T^t[u](x) = u(x+t)$. Then,
\begin{equation} \begin{aligned}
Q[T^t[u]](x) &= 
\int_{-\infty}^{\infty} k(x-y) g(T^t[u](y)) \, dy \\
&= \int_{-\infty}^{\infty} k(x-y) g(u(y+t)) \, dy
\end{aligned} \end{equation}
Let $z=y+t$.
\begin{equation} \begin{aligned}
Q[T^t[u]](x)
&= \int_{-\infty}^{\infty} k(x-z+t) g(u(z)) \, dz \\
&= Q[u](x+t) \\
&= T^t[Q[u]](x)
\end{aligned} \end{equation}

\item 
Suppose $\lim_{x\to\infty}u(x)=\ell$.

\end{proof}
